---
title: "Code Book"
author: "Rohit Padebettu"
date: "September 5, 2016"
output: pdf_document
---

## Steps in Analysis
1. Load the `dplyr`,`readr` and `tidyr` libraries
2. Read the X.txt, y.txt and subject.txt files in the "test" and "train" directories into respective tibbles
3. Read the features.txt file and activity_labels.txt file in the downloaded file directory into tibbles
4. Merge the train and test datasets for each of X,y and subject files using `bind_rows()` and `bind_cols()` from readr package
5. Grep the features.txt file to get an index for columns with `mean()` and `std()` occurring in them
6. Use the index to subset the columns of the tibble `X` and tibble `feature`
7. Append a column to the tibble `y` by using `inner_join` with the `activity_lables` tibble. This associates descriptive names with each Activity in `y`
8. Give appropriate descriptive names to the columns in `y`,`subject` and use the subsetted `features_filtered` tibble to name the columns in subsetted `X_filt`
9. Combine the `subject`, `y`, `X_filt` tibbles to create a new dataset `DataSetMergedNamed` which now has descriptive variable names and only mean and standard deviation attributes
10. Tidy the dataset by
    + Removing the `ActivityId` attribute
    + melting the data using `gather()` from `dplyr` package to create a dataset with only 4 columns `Subject`,`Activity`,`Feature`,`Value`
    + Group the data set into sub-groups based on Subject, Activity and Feature using `group_by` function
    + Calculate the `mean()` of each such sub_group using the `summarize` function in `dplyr` package
    + Name the column `Average`
11. The resulting dataset `DataTidy` is tidy as each column is a variable, each row is an observation and each observational unit is a value in the table. 
12. A summary of `DataTidy` is displayed and result is saved as `TidyData.txt` and `TidyData.csv`

## Brief Description of TidyData.txt

The TidyData set has 4 columns given below
```{r kable,echo=FALSE}
DataTidy<-readRDS("./TidyData.RDS")
names(DataTidy)

head(DataTidy,5)

summary(DataTidy)

```


**Subject** :   Is the ID of the person subject to the test. There were 30 volunteers in the original test

**Activity**:   Is the Activity the subject performed while measured. It is encoded as a factor of 6 levels, listed above

**Feature** :   Is the feature extracted from the original data set. There are 66 such features selected from the original dataset of 561 features.

**Average** :   Is the average value of a particular feature for the subject when performing a specific activity


## Feature Selection 
*(reproduced from original feature_info.txt document)*

*The features selected for this database come from the accelerometer and gyroscope 3-axial raw signals tAcc-XYZ and tGyro-XYZ. These time domain signals (prefix 't' to denote time) were captured at a constant rate of 50 Hz. Then they were filtered using a median filter and a 3rd order low pass Butterworth filter with a corner frequency of 20 Hz to remove noise. Similarly, the acceleration signal was then separated into body and gravity acceleration signals (tBodyAcc-XYZ and tGravityAcc-XYZ) using another low pass Butterworth filter with a corner frequency of 0.3 Hz.*

*Subsequently, the body linear acceleration and angular velocity were derived in time to obtain Jerk signals (tBodyAccJerk-XYZ and tBodyGyroJerk-XYZ). Also the magnitude of these three-dimensional signals were calculated using the Euclidean norm (tBodyAccMag, tGravityAccMag, tBodyAccJerkMag, tBodyGyroMag, tBodyGyroJerkMag).*

*Finally a Fast Fourier Transform (FFT) was applied to some of these signals producing fBodyAcc-XYZ, fBodyAccJerk-XYZ, fBodyGyro-XYZ, fBodyAccJerkMag, fBodyGyroMag, fBodyGyroJerkMag. (Note the 'f' to indicate frequency domain signals).*

These signals were used to estimate variables of the feature vector for each pattern: 

```
tBodyAcc-XYZ
tGravityAcc-XYZ
tBodyAccJerk-XYZ
tBodyGyro-XYZ
tBodyGyroJerk-XYZ
tBodyAccMag
tGravityAccMag
tBodyAccJerkMag
tBodyGyroMag
tBodyGyroJerkMag
fBodyAcc-XYZ
fBodyAccJerk-XYZ
fBodyGyro-XYZ
fBodyAccMag
fBodyAccJerkMag
fBodyGyroMag
fBodyGyroJerkMag
```

The set of variables that were estimated from these signals are: 

```
mean(): Mean value
std(): Standard deviation
```

## Feature List

The entire list of 66 features contained in this data set are as follows

```{r,echo=FALSE}
features <- read.table(file = "./UCI HAR Dataset/features.txt", header = FALSE)
meanIdx <- grep(pattern = "mean()",as.character(features$V2),value = FALSE,fixed = TRUE)
stdIdx <- grep(pattern = "std()",as.character(features$V2),value = FALSE,fixed = TRUE)
Idx <- sort(c(meanIdx,stdIdx),decreasing = FALSE)
features_filtered <- features[Idx,]
print(as.matrix(features_filtered$V2))
```

